Architectures for workflow integration in science gateways

### 2016-04-11 paper.tex
Updates in section 3 (first complete draft)

\usepackage{svg}
\usepackage{url}
\usepackage{xspace}

@@

@@

@@

We define architectures from their main components and the interactions
between these components.

A workflow engine is not supposed to be aware of the science gateway (?).

@@

\subsection{Common components}

Infrastructure: computing and storage resources. Includes
  schedulers, catalogues, etc. Can be grid, cloud or even
  cluster. Some workflow engines may assume a shared file system
  usually available only on clusters.

  Workflow engine: executes workflows, i.e. process their dependencies
  and create computing tasks. Might transfer data too.

@@

a.

c. Data movements can be triggered by the user, to upload input data
or download processed data, or by the workflow engine, to transfer
data across the infrastructure.

d. Sub-tasking: science gateway knows the sub-tasks of a
task. Therefore, it can kill all sub-tasks of a task when a task is killed.

g. Workflow integration results in an interface, for instance a web form,
where users can enter the parameters of the workflow to be
executed. Integrating a workflow is not the same process as
integrating a workflow engine. Integrating a workflow engine results
in an interface that allows developers to enter the parameters of the
workflow engine.

Only workflow execution is described. 
Workflow integration is not covered.

@@

See Figure~\ref{archi:sub-task}. The science gateway is responsible to
execute the tasks on the infrastructure, dealing with heterogeneous
batch managers and meta-schedulers. The workflow engine is a
particular task that can submit sub-tasks to the science gateway. The
workflow engine keeps track of the dependencies between the sub-tasks
but their execution is delegated to the science gateway. The science
gateway may implement mechanisms to deal with task dependencies such
as basic dependency lists as available in most batch managers. It has
no global vision of the workflow. The fact that tasks are submitted
through the science gateway does not restrict the range of possible
execution models. Pilot jobs, for instance, can still be used.

@@

\subsection{Pool model}

See Figure~\ref{archi:agent}. Workflows are submitted by the science
gateway to a pool to which agents connect asynchronously to retrieve
and execute workflows (interaction \texttt{f}). Agents may be started according to various
policies, for instance to ensure load balancing. Agents may wrap
different types of workflow engines. This model was implemented in the
SHIWA pool~\cite{ROGE-13}.
\begin{figure}
\centering
\def\svgwidth{0.8\columnwidth}
\input{images/agent.pdf_tex}
\caption{Pool model}
\label{archi:agent}
\end{figure}

@@

See Figure~\ref{archi:import}.

= Examples
* IWIR (SHIWA FGI). Language conversion.
* first Niak+CBRAIN integration?

= Evaluation
* To interpret the language, you need either a generic interpreter or a conversion tool. Quite complex and hard to  maintain.

>>>>

\usepackage{svg}
\usepackage{url}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{hyperref}

@@

  Workflow engine: executes workflows, i.e. process their dependencies
  and create computing tasks. Might transfer data too. A workflow
  engine is not supposed to be aware of the science gateway.

@@

The infrastructure consists of the computing and storage resources
involved in the workflow execution, and the software services used to
access these resources. Infrastructures can be servers, clusters,
grids or clouds. Some workflow engines may assume specific
characteristics about the infrastructure, such as the presence of a
shared file system between the computing nodes.

@@

Architectures are defined below from their main components and the
interactions between these components.

@@

@@

The interactions described below are labeled consistently with the
notations used on Figures~\ref{archi:tight}-\ref{archi:import}.

\begin{enumerate}[leftmargin=0cm,itemindent=0.6cm,label=\texttt{(\alph*)}]

\item Workflow integration: consists in adding a new workflow to the
  science gateway so that users can execute it. It is conducted by
  science gateway administrators, developers or users, and it results
  in an interface, for instance a web form, where users can enter the
  parameters of the workflow to be executed. Integrating a workflow is
  not the same process as integrating a workflow engine.
\item Task control: operations to manage tasks on the infrastructure,
  including: submission, monitoring, termination, deletion, etc.
\item Data control: operations to manage data on
  the infrastructure, such as: upload, download, deletion, browsing,
  replication, caching, etc. Data movements can be triggered by the
  user, to upload input data or download processed data - interaction
  \texttt{(c1)}, or by the workflow engine, to transfer data across the
  infrastructure - interaction \texttt{(c2)}. \todo{add c1 and c2 to the
    Figures}
\item Workflow control: operations to execute a workflow, including:
  submission, monitoring, termination, etc. Workflow control can be
  coarse-grained (a.k.a. black box) or fine-grained (white box). In a
  coarse-grained model, workflow activities are masked. 
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc.
\item Pool-agent: specific to the architecture described in~\ref{sec:pool}.
\item Workflow conversion: translation from one workflow language to
  another one.
\end{enumerate}

@@

See Figure~\ref{archi:sub-task}. The science gateway is responsible to
execute the tasks on the infrastructure, dealing with heterogeneous
batch managers and meta-schedulers. The workflow engine is a
particular task that can submit sub-tasks to the science gateway. The
workflow engine keeps track of the dependencies between the sub-tasks
but their execution is delegated to the science gateway. The science
gateway may implement mechanisms to deal with task dependencies such
as basic dependency lists as available in most batch
managers. Although it has no global vision of the workflow, it can
keep track of the sub-tasks submitted by a given task, for instance to
be able to cancel them when the task is canceled. The fact that tasks
are submitted through the science gateway does not restrict the range
of possible execution models: pilot jobs, for instance, can still be
used.

@@

\subsection{Pool model}
\label{sec:pool}
See Figure~\ref{archi:agent}. Workflows are submitted by the science
gateway to a pool to which agents connect asynchronously to retrieve
and execute workflows (interaction \texttt{f}). Agents may be started according to various
policies, for instance to ensure load balancing. Agents may wrap
different types of workflow engines. This model was implemented in the
SHIWA pool~\cite{ROGE-13}.
\begin{figure}
\centering
\def\svgwidth{0.8\columnwidth}
\input{images/agent.pdf_tex}
\caption{Pool model}
\label{archi:agent}
\end{figure}

@@

See Figure~\ref{archi:import}. This is an abstract model that we
instantiate with the service invocation architecture for
consistency. Workflows are integrated in the science gateway through
format conversion from a native format to the science gateway
format. This was implemented in the SHIWA Simulation Platform through
the IWIR language that provided a common language for portability
across grid workflow
systems~\cite{plankensteiner-prodan-etal:2013}.
